### 2023-05-26 paper/main.tex
Branchful code

\documentclass[11pt]{revtex4}

\usepackage[colorlinks=true]{hyperref}

\begin{document}
\title{Magnetic Phase Transitions in an Aperiodic Monotile}
\author{Jack Dinsmore}
\affiliation{Department of Physics, Stanford University}

\begin{abstract}
  Abstract
\end{abstract}

\maketitle

Hello there.

\end{document}